---
title: "Assignment 3 Task 3"
author: "Matt Koller"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)

```

## Introduction

In this document, I will read in, wrangle, and explore data about urchins observed in five locations near Santa Barbara from 2008 through 2020. I will assess whether the data is normally distributed, and create a graph of the data.

## I. SECTION 1

## 1. Read in data

Here I will read in, clean, and wrangle the urchins data.

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% #Read in the urchins data from the data subfolder
  clean_names() #clean the column names to read in lower snake case
```

Here I will wrangle my data to show just the urchin name and size, so I can begin to play around with it.

```{r}
urchin_size <- urchins %>% #create a new dataset named urchin_size, comprised of...
  select(size,  common_name) #the size and common_name columns
```

## 2. Exploring data visually

I will now create a histogram for both Purple Urchins and Red Urchins, and explore whether the data appears to be normally distributed for both. 

Update: according to my histograms, the data appears to be normally distributed for both purple urchins and red urchins
```{r}
ggplot(data = urchin_size, aes(x = size)) + #create a ggplot based on the urchin_size data; the size will go along the x-axis
  geom_histogram() + #the ggplot will be a histogram, with bins equalling 20
  facet_wrap(~common_name) + #this will create two different histograms, one for each kind of urchin
  labs(x = "Size (cm)", #label the x-axis
       y = "Count", #label the y-axis
       title = "Size distribution of purple vs. red urchins") #label the graph
```

I will now create a boxplot for both Purple Urchins and Red Urchins.

Update: The distribution of both the purple and red urchins appears to be normal! This exercise rocks (so far)!

```{r}
ggplot(data = urchin_size, aes(common_name, size))+ #crete a ggplot comparing common_name and size
  geom_boxplot() + #make this ggplot a boxplot
  labs(x = "Common name of urchin", #label the x-axis
       y = "Urchin size (mm)", #label the y-axis
       title = "Size distribution of purple vs. red urchins") #label the graph
```

## 3. Exploring data with descriptive statistics

In this section, I will create a table of summary statistics that examines the urchins dataframe.

```{r}
urchin_summary <- urchin_size %>%  #create a dataset from urchin_size (which only displays size and common name), that
  group_by(common_name) %>% #assesses the summary statistics broken down by common_name
  summarize(mean_size = mean(size), #assesses the mean size for both species
            median_size = median(size), #assesses the median size for both species
            std_dev = sd(size), #assesses the standard deviation for both species
            sample_size=n()) #provides the sample size for both species
       kbl(urchin_summary) %>% #this uses the kableExtra package to make the summary stats table look neat and tidy!
         kable_styling
```

## 4. Using words to interpret my summary data 

The histograms comparing the size of the purple and red urchins both appear to be normally distributed. The red urchins appear to be, on the whole, larger than the purple urchins. However, the purple urchins appear to be concentrated more closely around their median value, with less variance than the red urchins. The summary statistics table makes it clear that the red urchins have a mean size of 6.78 cm, while the purple urchins have a mean size of 4.47 cm, a difference of 2.31cm. The mean purple urchin size is 51.68% larger than the mean red urchin size. The percentage difference between the two is 41.07%.

```{r}
mean_size_purple_urchin <- 6.78 #Coding in the mean size for purple urchins
mean_size_red_urchin <- 4.47 #Coding in the mean size for red urchins
mean_difference <- mean_size_purple_urchin-mean_size_red_urchin #Calculating the difference in means
percentage_greater <- (mean_difference/mean_size_red_urchin)*100 #Calculating the percentage difference according to the following: on average, how much bigger (percentage-wise) are purple urchins than red urchins?
pct_dif <- function(mean_size_purple_urchin, mean_size_red_urchin){(mean_size_red_urchin-mean_size_purple_urchin)/((mean_size_purple_urchin+mean_size_red_urchin)/2)} #Use a function to calculate percentage difference between the two
pct_dif(mean_size_purple_urchin, mean_size_red_urchin) #calculate percentage difference

```


## II. SECTION 2

## 1. Wrangle data in a whole new way!

In this section, I will create a new data subset called purple_urchins. In this subset, I will wrangle the data by making the dates more clearly readable, and creating a new column detailing the site names.

```{r}
purple_urchins <- urchins %>% #this creates a new data subset called purple_urchins
  filter(common_name == "Purple Urchin") %>% #filtering only variables whose common_name is "Purple Urchin"
  mutate(date_new = mdy(date)) %>% #creates a new column displaying the data in "Date" class as YYYY-MM-DD
  mutate(year = year(date_new)) %>% #creates a new column called "year" displaying just the year of the variable
  mutate(site_full = case_when( #add a column that expands the abbreviated name of the site into the full site..
    site == c("NAPL") ~ "Naples Reef", #based on these parameters!
    site == c("AQUE") ~ "Arroyo Quemado Reef",
    site == c("CARP") ~ "Carpinteria Reef",
    site == c("MOHK") ~ "Mohawk Reef",
    site == c("IVEE") ~ "Isla Vista Reef"
    ))
```

## 2. Create a summary table

In this section, I will create a summary table containing the total number of purple urchins observed by site and year.

```{r}
purple_urchin_summary <- purple_urchins %>% #This creates a new dataset based on purple_urchins, called purple_urchin_summary...
  group_by(site_full, year) %>% #which creates two columns, grouping by the full site name and year...
  count(site_full) #and showing the number of variables counted grouped by both site name and year
```

## 3. Create an exploratory graph

In this section, I will create an exploratory graph assessing the number of urchins observed each year at five different locations.

```{r}

ggplot(data = purple_urchin_summary, aes(x = year, y = n))+ #this makes a graph comparing the year with the urchin count during each year. The "n" comes from the column generated in the previous code chunk by using the count function to tally the number of urchin observations at each site for each year.
  geom_col(aes(color = site_full)) + #This creates a bar graph demonstrating the counts of urchins by year at each reef. I originally tried doing a histogram and was getting tripped up, because the histogram only requires a single x variable and I couldn't find a way to give it a y variable (or somehow change my code in the previous chunk such that it was counting the number of observations), so I checked the Slack and others were having issues too....issues which were solved by using a geom_col!
  theme(legend.position="FALSE") + #removing legend from the graph
  facet_wrap(~site_full) #This creates a separate graph for each observation site!

```

## END

Success! This went really well!
